Run-Time Monitoring with Adjustable Overheads Using Dataflow-Guided Filtering

### 2014-09-12 dropping.tex
Minor edits from Mohamed's comments.

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as check
found for line 4 in Figure~\ref{fig:dropping.example_full} can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to add a 1-bit flag to each metadata in order to mark these
metadata as
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

In order to efficiently support dropping monitoring events and to prevent false
positives, we propose to insert a hardware module between the main and
monitor (see Figure~\ref{fig:dropping.dataflow_overview}). 
This module handles the invalidation operations shown in the middle column of
Figure~\ref{fig:dropping.example_invalid}.
There are two operations that are done for handling invalidation information:

@@

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
metadata. Null metadata corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes.

>>>>

The key cause of false positives is dropping monitoring operations that update
metadata. Dropping monitoring operations that check for an error (such as the check
for line 4 in Figure~\ref{fig:dropping.example_full}) can only cause
false negatives but will never cause false positives. On the other hand,
skipping monitoring operations that update metadata can lead to false positives
and false negatives.
Essentially, when an update operation is skipped, we can no longer trust the
corresponding metadata. Thus, our general approach for preventing false
positives is to add a 1-bit flag to each metadata in order to mark these
metadata as
invalid. Furthermore, this invalidation information is propagated in the same
way that metadata is. Figure~\ref{fig:dropping.example_invalid}
shows an example of associating invalid flags with metadata. In the example, no
monitoring operations are dropped. However, suppose that the monitoring for
line 1 is dropped. In this case, {\tt rf\_invalid[1]} will be marked as {\tt
true}. Thus, when line 3 is reached, the monitoring event is dropped
since {\tt rf\_invalid[1]} is marked as true. Note that in this case, line 2
also propagates this invalid flag to {\tt rf\_invalid[2]} and causes the check
performed on line 4 to also be dropped. This is necessary because an error
would have been raised even if the access on line 4 was in bounds.

@@

In order to efficiently support dropping monitoring events and to prevent false
positives, we propose to insert a hardware module between the main core and
monitor (see Figure~\ref{fig:dropping.dataflow_overview}). 
This module handles the invalidation operations shown in the middle column of
Figure~\ref{fig:dropping.example_invalid}.
There are two operations that are done for handling invalidation information:

@@

One way to reduce the number of monitoring events that must be handled by the
monitor is to filter out monitoring events that correspond to operating on null
metadata. Null metadata corresponds to events that
are not relevant to the monitor. For example, Hardbound
\cite{hardbound-asplos08} filters out operations on non-pointer (i.e., no base
and bounds metadata) instructions since it is not relevant to array bounds
checking. More recently, FADE \cite{fade-hpca14} has been proposed as a general
hardware module to perform this null metadata filtering for a variety of
monitoring schemes.

### 2014-09-12 evaluation.tex
Minor edits from Mohamed's comments.

In addition to evaluating our system for a core-based monitor, we also
evaluated partial monitoring on a higher performance FPGA-based monitor. This
setup is based on the FlexCore \cite{flexcore-micro10} setup and uses a
fully-pipelined monitor running on an FPGA fabric at half the frequency of the
main core. We show results for UMC due to space constraints though other
monitors show similar trends. In this case, full overheads of
monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
the coverage achieved as we sweep overhead from 1.01-1.7x. We see that partial
monitoring can allow the overheads of such a system to be pushed below 10\%
while still providing over 50\% coverage.

>>>>

In addition to evaluating our system for a core-based monitor, we also
evaluated partial monitoring on a higher performance FPGA-based monitor. This
setup is based on the FlexCore \cite{flexcore-micro10} setup and uses a
fully-pipelined monitor running on an FPGA fabric at half the frequency of the
main core. We show results for UMC due to space constraints though other
monitors show similar trends. In this case, the full overheads of
monitoring range from 1.1-1.7x. Figure~\ref{fig:evaluation.fpga_umc_sweep} shows
the coverage achieved as we sweep the overhead target from 1.01-1.7x. We see that partial
monitoring can allow the overheads of such a system to be pushed below 10\%
while still providing over 50\% coverage.

### 2014-09-12 monitoring.tex
Minor edits from Mohamed's comments.

There have been a number of proposals for run-time monitoring systems exploring various
design points. % in the trade-off space between efficiency and flexibility.
Table~\ref{tab:monitoring.previous_overheads} summarizes some of representative designs
and their reported performance overhead. The previous studies clearly show that there
exist trade-offs among efficiency, flexibility, and hardware costs. 
For example, a run-time monitoring scheme can often be realized with fairly low
performance overhead (less than 20\%) if implemented with custom hardware that is
designed only for one monitor or a narrow set of monitors. However, the custom
hardware monitors cannot be modified or updated, and require dedicated silicon area. 
On the other hand, flexible systems that support a wide range of monitors lead 
to noticeable performance overhead, often too high for wide deployment in practice.
Software-only implementations \cite{nagarajan-interact08, lift-micro06,
purify-usenix92, taintcheck-ndsss05} or multi-core monitors with minimal
hardware changes \cite{lba-asid06} are reported to have severalfold slowdowns.
On-chip FPGA monitors \cite{flexcore-micro10} and cores with monitoring accelerators
\cite{lba-isca08, fade-hpca14} can reduce overhead significantly, but still show
slowdowns of several tens of percents in some cases.
In today's monitoring systems, the overhead is also unpredictable because they
depend heavily on the characteristics of applications and monitoring operations.
In summary, users currently need to either pay noticeable overhead or the cost of custom
hardware in order to use fine-grained run-time monitoring in deployed systems.

@@

{\bf Profiling:} 
The run-time monitoring system can be used to implement various profiling tools
to collect statistics on program behaviors for performance optimizations as well
as security protection. For example, a recent study showed that an instruction
mix can be used to identify malware from normal programs \cite{demme-isca13, tang-raid14}. 
In such profiling tools, the partial monitoring framework can be used to obtain
the statics on samples rather than all program events, essentially trading off
accuracy for lower overhead.

@@

\begin{enumerate}
  \item \textbf{General:} Because we mainly target flexible run-time monitoring systems,
  which often have high overhead, the framework also needs to be general enough to be
  applicable to a wide range of monitoring scheme.

  \item \textbf{No false positive:} The framework needs to ensure that dropping a portion
  of monitoring operations does not lead to a false positive. We found that a data-flow
  that tracks invalid meta-data can serve as a general solution to this problem 
  (Section~\ref{sec:dropping}).

>>>>

There have been a number of proposals for run-time monitoring systems exploring various
design points. % in the trade-off space between efficiency and flexibility.
Table~\ref{tab:monitoring.previous_overheads} summarizes some the of representative designs
and their reported performance overhead. The previous studies clearly show that there
exist trade-offs between efficiency, flexibility, and hardware costs. 
For example, a run-time monitoring scheme can often be realized with fairly low
performance overhead (less than 20\%) if implemented with custom hardware that is
designed only for one monitor or a narrow set of monitors. However, the custom
hardware monitors cannot be modified or updated, and require dedicated silicon area. 
On the other hand, flexible systems that support a wide range of monitors lead 
to noticeable performance overhead, often too high for wide deployment in practice.
Software-only implementations \cite{nagarajan-interact08, lift-micro06,
purify-usenix92, taintcheck-ndsss05} or multi-core monitors with minimal
hardware changes \cite{lba-asid06} are reported to have severalfold slowdowns.
On-chip FPGA monitors \cite{flexcore-micro10} and cores with monitoring accelerators
\cite{lba-isca08, fade-hpca14} can reduce overhead significantly, but still show
slowdowns of several tens of percents in some cases.
In today's monitoring systems, the overhead is also unpredictable because they
depend heavily on the characteristics of applications and monitoring operations.
In summary, users currently need to either pay noticeable overhead or the cost of custom
hardware in order to use fine-grained run-time monitoring in deployed systems.

@@

{\bf Profiling:} 
The run-time monitoring system can be used to implement various profiling tools
to collect statistics on program behaviors for performance optimizations as well
as security protection. For example, a recent study showed that an instruction
mix can be used to identify malware from normal programs \cite{demme-isca13, tang-raid14}. 
In such profiling tools, the partial monitoring framework can be used to obtain
statistical samples rather than complete counts of all program events, essentially trading off
accuracy for lower overhead.

@@

\begin{enumerate}
  \item \textbf{General:} Because we mainly target flexible run-time monitoring systems,
  which often have high overhead, the framework also needs to be general enough to be
  applicable to a wide range of monitoring schemes.

  \item \textbf{No false positive:} The framework needs to ensure that dropping a portion
  of monitoring operations does not lead to a false positive. We found that a dataflow engine
  that tracks invalid metadata can serve as a general solution to this problem 
  (Section~\ref{sec:dropping}).
